Implementing Capon Beamforming on a GPU for Real-Time Cardiac Ultrasound Imaging

### 2013-03-26 HRUIonGPU/jrnl_article.tex
Finished tuffc article on gpu capon

Previous work on parallel implementations of the Capon beamformer have been focused on radar and passive sonar systems using custom made systolic array processors \cite{McWhirter1989, Moonen1993, Sinha2002}. For medical ultrasound, Chen et al. \cite{Chen2011a, Chen2011, Chen} implemented a Capon beamformer on the GPU that produces real apodization weights based on real data. However, the Capon beamformer is meant to process complex data. Under the constraint of using real data, the Capon beamformer is not allowed to micro-steer the main lobe, and therefore its full potential is not reached. Micro-steering of the main lobe is an important property of the Capon beamformer in order for it to improve structural details and edge definitions in an ultrasound image \cite[Fig~9.]{Synnevag2009}. Complex arithmetic leads to more operations per sample, but also allows the sampling frequency to be reduced. The number of samples needed per image can therefore be minimized. All results in this paper are reported with the use of baseband (or I/Q) data, where the sampling frequency is near the system bandwidth.

@@

In the next section we present background information on Capon beamforming in both element-space and beamspace and the GPU computing model. In Section \ref{sec:meth} we recap our parallel implementation of the Capon beamformer, before the beamspace method, and its GPU implementation, is presented in Section \ref{sec:bs}. Benchmark results and resulting images are presented in Section \ref{sec:bench} and \ref{sec:images}, together with a discussion of trading image quality for speed in Section \ref{sec:trade}.  Finally we discuss the results in Section \ref{sec:dis}, and draw conclusions in Section \ref{sec:con}.

@@

As shown in Fig. \ref{fig:mvbf}, the basic idea of an adaptive beamformer is to form a complex weight vector based on the received data.   The Capon beamformer produces weights that minimize the beamformer output power while maintaining unity gain in the focus direction. The result is that interference impinging from other directions is suppressed \cite{Synnevag2007}. Micro-steering caused by the complex weights is minute compared to the initial steering and focusing step, therefore the beamformer also works well for broadband ultrasound signals.

@@

The Capon beamformer, following the approach in \cite{Synnevag2009}, can be divided into three main steps: estimation of a sample covariance matrix, solving a system of linear equations, and calculation of the beamformer output.  The sample covariance matrix is estimated as 
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,\label{eq:R}
\end{align}
where  $N_K = 2K + 1$ is the number of temporal samples included, $\vec{x}_l = [x_l[n], \dotso, x_{l+L-1}[n]]$ is the $l\text{th}$ subarray of length $L$, and $N_L = M-L+1$ is the number of subarrays. Note that the covariance matrix dimension is $L \times L$. The covariance matrix is loaded with a diagonal factor $\epsilon$ to ensure numerical stability and increased robustness, 
\begin{align}\label{eq:diag}
\mat{\hat{R}}[n] = \mat{\breve{R}}[n] + \epsilon[n]\mat{I} = f(\vec{x}[n]).
\end{align}
Where $f(\cdot)$ represents the process of constructing $\mat{\hat{R}}$ from the input $\vec{x}$.
A factor proportional to the output power is often used to reduce the need for parameter adjustments, and the trace of $\mat{\breve{R}}$, 
\begin{align}\label{eq:diag_adapt}
\epsilon[n] &= d \times \frac{\text{trace}\{\mat{\breve{R}}[n]\}}{L},
\end{align}
has been applied in much of the recent literature on Capon beamforming for medical ultrasound imaging \cite{Synnevag2007, Nilsen2009, Wang2009, Mehdizadeh2012}. It has also been applied to HF and VHF antenna processing \cite{Featherstone1997b}.

@@

In this section we give a brief discussion of GPU computing, laying the basis for the discussions on parallel implementation strategies. The Capon beamformers have been implemented using the CUDA GPU framework by Nvidia \cite{Nvidia2013}, however the challenges and proposed methods described in this paper are valid for other GPU framworks as well, like e.g. OpenCL. Independent of framework, the key point is that GPUs can accelerate large sets of small concurrent tasks, like beamforming a pixel in a large image, much better than a CPU. While the CPU has few cores and lack logic to process complex threads, the GPU has many cores, but each core needs many (preferably simple) threads to keep it occupied and to thereby achieve maximum throughput.

@@

An SM has a limited amount of shared memory (or near-core cache) and registers. Therefore, to achieve maximum GPU occupancy (defined as having the maximum number of threads residing on each SM), each thread has to consume as few registers and as little shared memory as possible. Each thread should also perform enough instructions per transferred byte in order to hide memory latency. Dividing the problem into enough threads will help to hide instruction latency. Global GPU memory transactions, in addition to CPU-GPU transfers, are slow and must be minimized. All of these constraints combined pose a real challenge as the target problem has to map well to the GPU framework on a grid, block, thread, and memory consumption level in order to benefit from parallel acceleration. In the next section, we will describe how this mapping can be done for the Capon beamforming algorithm.

@@

For the adaptive diagonal loading factor in (\ref{eq:diag}), the trace of $\mat{\hat{R}}$ has to be known before diagonal elements are written back to global memory. Since one row for all covariance matrices is written to global memory per kernel iteration, the diagonal elements need to be held in a shared memory until the trace is accumulated. When that is done, scattered writes of the scaled diagonal elements to global memory has to be performed.

@@

For the results presented in this paper, we have used a GPU implementation of batched Gauss-Jordan elimination made by Nvidia to solve $\mat{\hat{R}}\vec{b} = \vec{1}_L$, where $\vec{b} = \mat{\hat{R}}^{-1}\vec{1}_L$, for all pixels in one kernel launch. The solver is available online for registered CUDA developers \cite{Nvidiaa}.%For smaller systems, $L < 10$, a direct solver using one thread per system is preferable. For this purpose we have implemented Cramer's rule for $2 \times 2$, $3 \times 3$ and $4 \times 4$ matrices. All solvers have been benchmarked against the latest Intel MKL libraries.

@@

In order to reduce the complexity of the matrix inversion problem the transformation must be performed before or after the covariance matrix has been estimated. If the transformation is applied before, one transformation is needed per subarray. After this transformation the subarrays no longer overlap. The amount of input data are therefore changed from $M$ to $N_LN_b$ elements per sample, but all remaining steps will get lower computational complexity as long as $N_b < L$.

@@

As presented in (\ref{eq:diag}), adaptive diagonal loading is performed by loading the sample covariance matrix with the average energy per element, that is $\text{trace}\{\mat{\hat{R}}\}/L$. When data are transformed into beamspace, a high percentage of the total energy is mapped to a small subset of beams. Thus the average energy per beam in a reduced beamspace is, for a focused system, higher than the average energy per element. This means that less diagonal loading (smaller $d$) has to be applied when the covariance is estimated from beams instead of elements.

@@

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.45in]{gfx/benchmark_bar_bs_M=64_Nx=80_Ny=540_Nb=3.png} \label{fig:benchBS64}}
\hfil
\subfloat[]{\includegraphics[width=3.45in]{gfx/benchmark_bar_bs_M=96_Nx=120_Ny=540_Nb=3.png}%
\label{fig:benchBS96}}}
\caption{Benchmarks of GPU beamspace Capon beamforming for a cardiac image covering a $70 \degree$ and $15$ cm deep sector. Execution time is plotted as a function of subarray length $L$.  If $L$ had been kept constant on $M/2$ and execution time was plotted against $N_b$, the plot would be similar to the one in Fig. \ref{fig:bench}. The variation in execution time is caused by adaptive kernel launch configurations.}
\label{fig:benchBS}
\end{figure*}
In this section we present benchmarks of the implementations presented in Sections \ref{sec:meth} and \ref{sec:bs} performed on a Nvidia Quadro 6000 graphics card (14 SMs, 1 TFlops theoretical single precision performance, 6 GB global memory, and 144 GB/s memory bandwidth). Problem sizes are calculated based on 64 and 96 element phased arrays with a 3.4 MHz center frequency. Assuming a system bandwidth of 80\% (relative to the center frequency) we need a complex sampling frequency of 2.7 MHz, and when imaging down to a depth of 15.4 cm we have approximately 540 samples per range line. To sample a $70\degree$ sector with 64- and 96-element arrays ($d=\lambda/2$ pitch) at twice the Rayleigh criterion ($\lambda/(M*d)$), we need at approximately 80 and 120 receive beams respectively. The presented execution times do not include transfer of data from CPU- to the GPU-side, since data are assumed to be pre-delayed and available in global GPU memory. What is included are all calculations, memory transfers from global to shared GPU memory, and shared memory access.

@@

To investigate the image quality of both the ES and BS-Capon beamformer a phantom was simulated using Field II \cite{Jensen1992, Jensen1996a} (Fig. \ref{fig:phantom}). The simulation was performed using a 64-element, $\lambda/2$ pitch, 10 $\mu$m kerf, 80\% bandwidth, and 10 mm high phased array. The transducer was excited with a 3.4 MHz and $1.5\lambda$ long pulse with transmit focus at 90 mm. The phantom contains two circular areas, one bright at 20 dB and one dark at -20 dB, and two point targets at 30 dB, located at 90 mm depth. Speckle level is at 0 dB, covering an area from $\pm90\degree$ in azimuth. The resulting image has been processed both using DAS with uniform weights, ES and BS-Capon. The presented image is the first frame in a series of images where motion is imposed on both circles, point targets and the probe. The entire simulation is presented in three attached videos \multimedia{Media-Movie 1 and 2}.

Fig. \ref{fig:invivo} presents results of applying ES and BS-Capon on an \textit{in-vivo} image of the left ventricle. The image was acquired using a 64-element and $0.65\lambda$ pitched phased array in harmonic mode (1.7 MHz at transmit and 3.4 MHz at receive). The image is extracted from a cardiac cycle, that can be viewed in its full length in the attached videos \multimedia{Media-Movie 3, 4 and 5}. It is included to show that ES-Capon and BS-Capon beamforming are both feasible for cardiac imaging, not just in simulations.

@@

Both data sets have been processed with the following settings for the ES-Capon beamformer: $L = 32$, $K = 2$, and $d=0.1$. These settings have previously been shown to provide improved lateral resolution while maintaining DAS-like speckle \cite{Synnevag2007a}. For BS-Capon the following parameters have been used: $L=32$, $K=2$, $d=0.01$, and $N_b=3$. For the simulated phantom, 16x lateral oversampling on transmit has been performed in order to avoid under sampling of the two bright point scatterers as they move. The cardiac recording of a 70\degree{} sector has a minimal lateral sampling of 98 receive beams.

\section{Trading resolution for speed}\label{sec:trade}
As seen from Fig. \ref{fig:bench}, execution time is reduced when $L$ is reduced. In Fig. \ref{fig:benchBS} we see that execution time is also reduced when moving to beamspace. So to gain an  increase in processing throughput we can reduce $L$ in element space or $N_b$ in beamspace. In this section we investigate how adjusting $L$ for ES-Capon and $N_b$ for BS-Capon impacts the lateral resolution. In Fig. \ref{fig:speed_res_trade}, the minimum distance between the two point targets in Fig. \ref{fig:phantom} (while preserving a 6 dB saddle point in between the two) is plotted against $L$ (when processed with ES-Capon) and $N_b$ (when processed with BS-Capon). For the result in Fig. \ref{fig:speed_res_trade}, $d=0.01$ has been used on both methods in order to obtain the same resolution for $L=N_b=M/2$. This is considered a small value for ES-Capon and a normal value for BS-Capon (cf. Section \ref{sec:bs}), hence the resolution for ES-Capon is here better than what is presented in Fig. \ref{fig:phantom}. The precision of the presented measurements is 0.12 mm, dictated by the simulated frame rate, 25 frames per second (FPS), and the relative speed of the two point scatterers, 3 mm/s.

@@

In element space, resolution decays rapidly since the effective adaptive aperture is proportional to $L$. For $L = N_b=3$, BS-Capon provides almost 1 mm improved resolution over DAS and ES-Capon. However, note that the resolution of the Capon beamformers strongly depends on image SNR and the choice of parameters. Thus, different SNR scenarios and choices of parameters will lead to different results. It is also worth noting that ES-Capon beamforming in this example performs worse than DAS for small values of $L$. For these small subarrays, the amount of robustification is too high in order to correctly steer a zero at each point target or to accurately micro-steer the main lobe. Subarray averaging further impose a triangular or trapezoid apodization function, which leads to worse resolution if the introduced adaptivity can not compensate for the increase in main lobe width. For $L=1$ the ES-Capon is equal to DAS.

@@

\section{Discussion}\label{sec:dis}
In this paper we have presented a parallel implementation of the Capon beamformer in both element- and beamspace for high-resolution cardiac ultrasound imaging. The results show that it is now possible to apply BS-Capon in real-time in cardiac ultrasound imaging by utilizing the power of GPUs. According to Fig. \ref{fig:benchBS} it is also possible to process two parallel receive lines per transmit (for both $M=64$ and $M=96$. $L=M/2$). However, these extra lines might be needed for increased sampling instead of increased frame rate when using the Capon beamformers in high SNR situations (Fig \ref{fig:phantom}). If parallel receive lines are used to obtain denser sampling, acquisition time remains the same but processing times are increased with a factor equal to the amount of oversampling. If parallel receive lines are used for increased frame rate, the acquisition time is lowered, and the processing time remains constant. Hence, four parallel receive lines are not possible to process in real-time using BS-Capon (with $N_b=3$) in combination with our single GPU.

@@

Turning to image quality we see from Fig. \ref{fig:phantom}, and the attached video of the simulation, that Capon beamforming provides improved lateral resolution of moving point targets both in element- and in a reduced beam-space ($N_b=3$). Our preliminary finding is that the Capon beamformers are local shift invariant as long as the lateral sampling is increased to a curtain level \multimedia{Media-Movie 1 and 2}. How large this oversampling factor needs to be and how the oversampling should be conducted is outside the scope of this paper. No major improvement in contrast is visually observed in the presented images images and videos, other than slightly rounder circles due to mainlobe micro steering. In the cardiac recording (Fig. \ref{fig:invivo}), structures are thinner and better resolved when the Capon beamformers are applied. However we observe a loss in contrast, especially in the septum (Fig. \ref{fig:slices}). This loss can be partially explained by the presence of highly correlated signals and mismatch between the assumed steering vector and signal. Future work should focus on solving this issue.

@@

% use section* for acknowledgement
\section*{Acknowledgment}
The authors would like to thank Dr Kjell Kristoffersen, GE Vingmed Ultrasound, for valuable comments.

>>>>

Previous work on parallel implementations of the Capon beamformer have been focused on radar and passive sonar systems using custom made systolic array processors \cite{McWhirter1989, Moonen1993, Sinha2002}. For medical ultrasound, Chen \textit{et al.} \cite{Chen2011a, Chen2011, Chen} implemented a Capon beamformer on the GPU that produces real apodization weights based on real data. However, the Capon beamformer is meant to process complex data. Under the constraint of using real data, the Capon beamformer is not allowed to micro-steer the main lobe, and therefore its full potential is not reached. Micro-steering of the main lobe is an important property of the Capon beamformer in order for it to improve structural details and edge definitions in an ultrasound image \cite[Fig~9.]{Synnevag2009}. Complex arithmetic leads to more operations per sample, but also allows the sampling frequency to be reduced. The number of samples needed per image can therefore be minimized. All results in this paper are reported with the use of baseband (or I/Q) data, where the sampling frequency is near the system bandwidth.

@@

In the next section we present background information on Capon beamforming in both element-space and beamspace and the GPU computing model. In Section \ref{sec:meth} we recap our parallel implementation of the Capon beamformer, before the GPU implementation of the beamspace Capon beamformer is presented in Section \ref{sec:bs}. Benchmark results and resulting images are presented in Section \ref{sec:bench} and \ref{sec:images}, together with a discussion of trading image quality for speed in Section \ref{sec:trade}.  Finally we discuss the results in Section \ref{sec:dis}, and draw conclusions in Section \ref{sec:con}.

@@

As shown in Fig. \ref{fig:mvbf}, the basic idea of an adaptive beamformer is to form a complex weight vector based on the received data.   The Capon beamformer produces weights that minimize the beamformer output power while maintaining unity gain in the focus direction. The result is that interference impinging from other directions is suppressed \cite{Synnevag2007}. The complex weights cause a micro-steering which is minute compared to the initial steering and focusing step, therefore the beamformer also works well for broadband ultrasound signals.

@@

The Capon beamformer, following the approach in \cite{Synnevag2009}, can be divided into three main steps: estimation of a sample covariance matrix, solving a system of linear equations, and calculation of the beamformer output.  The sample covariance matrix is estimated as 
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,\label{eq:R}
\end{align}
where  $N_K = 2K + 1$ is the number of temporal samples included, $\vec{x}_l = [x_l[n], \dotso, x_{l+L-1}[n]]$ is the $l\text{th}$ subarray of length $L$, and $N_L = M-L+1$ is the number of subarrays. Note that the covariance matrix dimension is $L \times L$. The covariance matrix is loaded with a diagonal factor $\epsilon$ to ensure numerical stability and increased robustness: 
\begin{align}\label{eq:diag}
\mat{\hat{R}}[n] = \mat{\breve{R}}[n] + \epsilon[n]\mat{I} = f(\vec{x}[n]).
\end{align}
The operator $f(\cdot)$ represents the process of estimating $\mat{\hat{R}}$ from the input $\vec{x}$.
A factor proportional to the output power is often used to reduce the need for parameter adjustments, and the trace of $\mat{\breve{R}}$, 
\begin{align}\label{eq:diag_adapt}
\epsilon[n] &= d \times \frac{\text{trace}\{\mat{\breve{R}}[n]\}}{L},
\end{align}
has been applied in much of the recent literature on Capon beamforming for medical ultrasound imaging \cite{Synnevag2007, Nilsen2009, Wang2009, Mehdizadeh2012}. It has also been applied to HF and VHF antenna processing \cite{Featherstone1997b}.

@@

In this section we give a brief discussion of GPU computing, laying the basis for our discussions on parallel implementation strategies. The Capon beamformers have been implemented using the CUDA GPU framework by Nvidia \cite{Nvidia2013}, however the challenges and proposed methods described in this paper are valid for other GPU framworks as well, like e.g. OpenCL. Independent of framework, the key point is that GPUs can accelerate large sets of small concurrent tasks, like beamforming a pixel in a large image, much better than a CPU. While the CPU has few cores, but logic to process complex threads, the GPU has many cores, but each core needs many (preferably simple) threads to keep it occupied and to achieve maximum throughput.

@@

An SM has a limited amount of shared memory (or near-core cache) and registers. Therefore, to achieve maximum GPU occupancy (defined as having the maximum number of threads residing on each SM), a conservative use of registers and shared memory per thread is required. Each thread should perform enough instructions per transferred byte in order to hide memory latency, and the problem should be divided into enough threads in order to hide instruction latency. Global GPU memory transactions, in addition to CPU-GPU transfers, are slow and must be minimized. All of these constraints combined pose a real challenge as the target problem has to map well to the GPU framework on both a grid, block, thread, and memory consumption level in order to benefit from parallel acceleration. In the next section, we will describe how this mapping can be done for the ES-Capon beamformer.

@@

For the adaptive diagonal loading factor in (\ref{eq:diag}), the trace of $\mat{\hat{R}}$ has to be known before diagonal elements are written back to global memory. Since one row for all covariance matrices is written to global memory per kernel iteration, the diagonal elements need to be held in shared memory until the trace is accumulated. When that is done, scattered writes of the scaled diagonal elements to global memory has to be performed.

@@

For the results presented in this paper, we have used a GPU implementation of batched Gauss-Jordan elimination made by Nvidia to solve $\mat{\hat{R}}\vec{b} = \vec{1}_L$, where $\vec{b} = \mat{\hat{R}}^{-1}\vec{1}_L$, for all pixels in an image in one kernel launch. The solver is available online for registered CUDA developers \cite{Nvidiaa}.%For smaller systems, $L < 10$, a direct solver using one thread per system is preferable. For this purpose we have implemented Cramer's rule for $2 \times 2$, $3 \times 3$ and $4 \times 4$ matrices. All solvers have been benchmarked against the latest Intel MKL libraries.

@@

In order to reduce the complexity of the matrix inversion problem the transformation must be performed before or after the covariance matrix has been estimated. If the transformation is applied before, one transformation is needed per subarray. This results in subarrays that no longer overlap. The amount of input data are therefore changed from $M$ to $N_LN_b$ elements per sample, but all remaining steps will get lower computational complexity as long as $N_b < L$.

@@

As presented in (\ref{eq:diag}), adaptive diagonal loading is performed by loading the sample covariance matrix with the average energy per element, that is $\text{trace}\{\mat{\breve{R}}\}/L$. When data are transformed into beamspace, a high percentage of the total energy is mapped to a small subset of beams. Therefore, if a set of high energy beams is selected, the average energy per beam will be higher than the average energy per element ($\text{trace}\{\mat{\breve{R}}\}/L < \text{trace}\{\mat{\breve{R}}_{\text{BS}}\}/N_b$). This means that less diagonal loading (smaller $d$) has to be applied when the covariance is estimated among beams and not elements.

@@

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.45in]{gfx/benchmark_bar_bs_M=64_Nx=80_Ny=540_Nb=3.png} \label{fig:benchBS64}}
\hfil
\subfloat[]{\includegraphics[width=3.45in]{gfx/benchmark_bar_bs_M=96_Nx=120_Ny=540_Nb=3.png}%
\label{fig:benchBS96}}}
\caption{Benchmarks of GPU beamspace Capon beamforming for a cardiac image covering a $70 \degree$ and $15$ cm deep sector. Execution time is plotted as a function of subarray length $L$.  If $L$ had been kept constant on $M/2$ and execution time was plotted against $N_b$, the plot would be similar to the one in Fig. \ref{fig:bench}. The variation in execution time is caused by adaptive kernel launch configurations.}
\label{fig:benchBS}
\end{figure*}
In this section we present benchmarks of the implementations presented in Sections \ref{sec:meth} and \ref{sec:bs} performed on a Nvidia Quadro 6000 graphics card (14 SMs, 1 TFlops theoretical single precision performance, 6 GB global memory, and 144 GB/s memory bandwidth). Problem sizes are calculated based on 64 and 96 element phased arrays with a 3.4 MHz center frequency. Assuming a system bandwidth of 80\% (relative to the center frequency) we need a complex sampling frequency of 2.7 MHz, and when imaging down to a depth of 15.4 cm we have approximately 540 samples per range line. To sample a $70\degree$ sector with 64- and 96-element arrays ($d=\lambda/2$ pitch) at twice the Rayleigh criterion ($\lambda/(2Md)$), we need approximately 80 and 120 receive beams respectively. The presented execution times do not include transfer of data from CPU- to the GPU-side, since data are assumed to be pre-delayed and available in global GPU memory. What is included are all calculations, memory transfers from global to shared GPU memory, and shared memory access.

@@

To investigate the image quality of both the ES and BS-Capon beamformer, a phantom was simulated using Field II \cite{Jensen1992, Jensen1996a} (Fig. \ref{fig:phantom}). The simulation was performed using a 64-element, $\lambda/2$ pitch, 10 $\mu$m kerf, 80\% bandwidth, and 10 mm high phased array. The transducer was excited with a 3.4 MHz and $1.5\lambda$ long pulse with transmit focus at 90 mm. The phantom contains two circular areas, one bright at 20 dB and one dark at -20 dB, and two point targets at 30 dB, located at 90 mm depth. Speckle level is at 0 dB, covering an area from $\pm90\degree$ in azimuth. The resulting image has been processed both using DAS with uniform weights, ES- and BS-Capon. The presented image is the first frame in a series of images where motion is imposed on both circles, point targets and the probe. The entire simulation is presented in two attached videos \multimedia{Media-Movie 1 and 2}.

Fig. \ref{fig:invivo} presents results of applying ES- and BS-Capon on an \textit{in-vivo} image of the left ventricle. The image was acquired using a 64-element and $0.65\lambda$ pitched phased array in harmonic mode (1.7 MHz at transmit and 3.4 MHz at receive). The image is extracted from a cardiac cycle, that can be viewed in its full length in the attached videos \multimedia{Media-Movie 3, 4 and 5}. It is included to show that both ES-Capon and BS-Capon beamforming are feasible for \textit{in-vivo} cardiac imaging, not just in simulations.

@@

Both data sets have been processed with the following settings for the ES-Capon beamformer: $L = 32$, $K = 2$, and $d=0.1$. These settings have previously been shown to provide improved lateral resolution while maintaining DAS-like speckle \cite{Synnevag2007a}. For BS-Capon the following parameters have been used: $L=32$, $K=2$, $d=0.01$, and $N_b=3$. For the simulated phantom, 16x lateral oversampling on transmit has been performed in order to avoid under sampling of the two bright point scatterers as they move. The cardiac recording of a 70\degree{} sector has a minimal lateral sampling of 98 receive beams ($\lambda/2Md$).

\section{Trading resolution for speed}\label{sec:trade}
As seen from Fig. \ref{fig:bench}, execution time is reduced when $L$ is reduced. In Fig. \ref{fig:benchBS} we see that execution time is also reduced when moving to beamspace. So to gain an  increase in processing throughput we can reduce $L$ in element space or $N_b$ in beamspace. In this section we investigate how adjusting $L$ for ES-Capon and $N_b$ for BS-Capon impacts the lateral resolution. In Fig. \ref{fig:speed_res_trade}, the minimum distance between the two point targets in Fig. \ref{fig:phantom} (while preserving a 6 dB saddle point in between the two) is plotted against $L$ (when processed with ES-Capon) and $N_b$ (when processed with BS-Capon). To produce the result in Fig. \ref{fig:speed_res_trade}, $d=0.01$ has been used on both methods in order to obtain the same resolution for $L=N_b=M/2$. We considered this to be a small value for ES-Capon and a normal value for BS-Capon (cf. Section \ref{sec:bs}), hence the resolution for ES-Capon is here better than what is presented in Fig. \ref{fig:phantom}. The precision of the presented measurements is 0.12 mm, dictated by the simulated frame rate, 25 frames per second (FPS), and the relative speed of the two point scatterers, 3 mm/s.

@@

In element space, resolution decays rapidly since the effective adaptive aperture is proportional to $L$. For $L = N_b=3$, BS-Capon provides almost 1 mm improved resolution over DAS and ES-Capon. However, note that the resolution of the Capon beamformers strongly depends on image SNR and the choice of parameters. Thus, different SNR scenarios and choices of parameters will lead to different results. It is also worth noting that ES-Capon beamforming in this example performs worse than DAS for small values of $L$. For these small subarrays, the method is not able to steer zeros at each point target or accurately micro-steer the main lobe. Subarray averaging further impose a triangular or trapezoid apodization function, which leads to worse resolution if the introduced adaptivity can not compensate for the increase in main lobe width. For $L=1$ the ES-Capon is equal to DAS.

@@

\section{Discussion}\label{sec:dis}
In this paper we have presented a parallel implementation of the Capon beamformer in both element- and beamspace for high-resolution cardiac ultrasound imaging. The results show that it is now possible to apply BS-Capon in real-time in cardiac ultrasound imaging by utilizing the power of GPUs. According to Fig. \ref{fig:benchBS} it is also possible to process two parallel receive lines per transmit beam (for both $M=64$ and $M=96$. $L=M/2$). These extra lines might be needed for increased sampling instead of increased frame rate when using the Capon beamformers in high SNR situations (Fig \ref{fig:phantom}). If parallel receive lines are used to obtain denser sampling, acquisition time remains the same but processing times are increased with a factor equal to the amount of oversampling. If parallel receive lines are used for increased frame rate, the acquisition time is lowered, and the processing time remains constant. Hence, four parallel receive lines are not possible to process in real-time using BS-Capon (with $N_b=3$) in combination with our single GPU setup.

@@

Turning to image quality we see from Fig. \ref{fig:phantom}, and the attached video of the simulation, that Capon beamforming provides improved lateral resolution of moving point targets both in element- and in a reduced beam-space ($N_b=3$). Our preliminary finding is that the Capon beamformers are local shift invariant as long as the lateral sampling is increased to a curtain level \multimedia{Media-Movie 1 and 2}. How large this oversampling factor needs to be and how the oversampling should be conducted is outside the scope of this paper. No major improvement in contrast is visually observed in the presented images and videos, other than slightly rounder circles due to mainlobe micro steering. In the cardiac recording (Fig. \ref{fig:invivo}), structures are thinner and better resolved when the Capon beamformers are applied. However we observe a loss in contrast, especially in the septum (Fig. \ref{fig:slices}). This loss can be partially explained by the presence of highly correlated signals and mismatch between the assumed steering vector and signal. Future work should focus on solving this issue.

@@

% use section* for acknowledgement
\section*{Acknowledgment}
The authors would like to thank Dr K. Kristoffersen, GE Vingmed Ultrasound, for valuable comments.
